---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
  keep_md: yes
  pdf_document: default
---


## Loading and preprocessing the data
```{r}
d <- read.csv('activity.csv')
```


## What is mean total number of steps taken per day?
```{r}
library(plyr)
out <- ddply(d, c("date"),summarize, steps=sum(steps))
n <- nrow(out)
mean_val <- sum(out$steps, na.rm = T) / n
median_val <- median(out$steps, na.rm=T)

mean_val
median_val 
```


```{r scatterplot0,fig.height=4}
hist(out$steps ,breaks=100, main=" total number of steps taken per day",
     xlab="steps", ylab="frequency", col="pink")
dev.copy(png,'plot0.png')
dev.off()
```


## What is the average daily activity pattern?

```{r}
days <- nrow(out)
interval_cnt <- 24*60/5   

out2 <- rep(0, interval_cnt)

for(i in 1:interval_cnt)
{
    some <- seq(i, nrow(d), by= interval_cnt)
    out2[i] <- mean( d[some,]$steps, na.rm=T)
}
```


```{r scatterplot1,fig.height=4}
plot(out2 , type="l", main="average daily activity pattern",
     xlab="5 minutes interval",ylab="steps", col="red" , lwd=3)
dev.copy(png,'plot1.png')
dev.off()
```



```{r}
max_interal <- max(out2)
```

## Imputing missing values
```{r}
d2 <- d
for(i in 1:nrow(d2) )
{
    if( is.na(d2[i,1])  == TRUE)
    {
       d2[i,1] <- out2[ i %% 288 + 1]

    }
}

out_imputing <- ddply(d2, c("date"),summarize, steps=sum(steps))
```


```{r scatterplot2,fig.height=4}
hist(out_imputing$steps ,breaks=100, 
     main="total number of steps taken per day after IMPUTING" , xlab="steps",ylab="frequency", col="magenta")
dev.copy(png,'plot2.png')
dev.off()
```

## Are there differences in activity patterns between weekdays and weekends?
As we expected, peopel tend to move more on the weekend.
I'll calculate the average value of steps on the weekdays/weekends.


FYI, I set the language to english becuase I found out weekdays() returns  the value from this setting. some web browser might not support mine.

I also made the new column weekend.if the date is weekend, it gets 1. 
```{r}
out3 <- ddply(d2, c("date"),summarize, steps=sum(steps))

weekdays_n <- 0
weekend_n <- 0

weekdays_sum <-0
weekend_sum <- 0

Sys.setlocale("LC_TIME", "English")

for(i in 1:nrow(out3))
{
    if( weekdays( as.Date(out3[i,1])) == "Sunday" || 
        weekdays( as.Date(out3[i,1])) == "Saturday")
    {
       weekend_n <- weekend_n +1 
       weekend_sum <- weekend_sum + out3[i,2]
       out3$weekend <- 1
    
    }
    else
    {
       weekdays_n <- weekdays_n + 1
       weekdays_sum <- weekdays_sum + out3[i,2]
       out3$weekend <- 0
    
    }
}

# print average steps in weekend & weekdays
print( weekend_sum / weekend_n )
print( weekdays_sum / weekdays_n )
```


